---
title: "SRER Litter Bags"
author: "Taylor A Portman"
date: "2022-09-28"
output: html_document
---

## SRER litter bags T3: Mass loss

Grass:   
D= DICA, Digitaria californica  
E= ERLE, Eragrostis lehmanniana (Invasive)   

Plot:   
IO= Invasive Open (ERLE dominant)  
IC= Invasive Canopy (ERLE dominant)  
NO= Native Open (DICA dominant)  
NC= Native Canopy (DICA dominant)  

Mass Loss= Mass Initial - Mass Final  


```{r, include=FALSE}
library(tidyverse)
library(ggplot2)
setwd("/Volumes/GoogleDrive/My Drive/projects/ERLE_fungal_symbionts/R")
```

```{r, include=FALSE}
#Import data
T3MassLoss<-read.csv("data/T3LitterBags.csv")

T3MassLoss$ID<-T3MassLoss$LitterbagCode
T3MassLoss$MesquiteNum<-as.character(T3MassLoss$Mesquite.)

T3dat<- T3MassLoss%>%
 dplyr:: select(ID, Grass, Plot, MesquiteNum, MassLoss)%>%
  na.omit()
```


### Data quality checking: 

1. Compare mass loss between grass type, DICA and ERLE 
2. Compare mass loss between the three collection sites (MesquiteNum) 

Takeaways:  
1. No difference in overall mass loss between plant type 
2. Mesquite #1 and #2 are similar with a slight decrease in overall mass loss for mesquite #3 
```{r}
#1. Compare between plant type and look at overall spread of data
ggplot(T3dat)+
  geom_boxplot(aes(Grass, MassLoss))+
  geom_point(aes(Grass, MassLoss))

#2. Compare between plot and plant type
ggplot(T3dat, aes(MesquiteNum, MassLoss, fill=Grass))+
  geom_boxplot()
```

### Look at overall trends in mass loss between site and grass type

Takeaways: 
Treatments of interest are IC (invasive canopy), and NC (native canopy)   
IC: DICA > ERLE   
NC: ERLE > DICA   
```{r}
#Compare between treatment and grass
ggplot(T3dat,aes(Plot, MassLoss, fill= Grass))+
  geom_boxplot()
  #geom_point(aes(color=Grass))

#Look at all data seperated  by mesquite#
ggplot(T3dat,aes(Plot, MassLoss, color= Grass))+
  geom_point()+
  facet_wrap(~MesquiteNum)
```

### Quick look at IC and NC for statistical significance : t.test

Takeaway: 
Neither plot treatment causes significant differences in mass loss between grass species  
IC: p=0.2751  
NC: p=0.1081
```{r}
# IC: invasive canopy
T3.IC<- T3dat%>%
  dplyr::filter(Plot=="IC")

t.test(MassLoss~Grass, data=T3.IC)
#not sig

# NC: native canopy
T3.NC<- T3dat%>%
  dplyr::filter(Plot=="NC")

t.test(MassLoss~Grass, data=T3.NC)
#not sig


```


### ANOVA: looking at Mass loss as a function of Plot*Grass type 

Takeaway: 
Interaction effect of plot and grass type on Mass loss is marginally significant (p=0.0585)

```{r}
plot<-aov(MassLoss~ Plot*Grass, data=T3dat)
summary(plot)
```

### ANOVA: looking at Mass Loss as a function of plot, looking at each grass separately 

Takeaway: No plot treatment effect when filtering by grass species 
ERLE: p=0.11
DICA: p=0.276
```{r}
T3.ERLE<- T3dat %>%
  dplyr::filter(Grass== "E")
T3.DICA<- T3dat %>%
  dplyr::filter(Grass== "D")

plot.grass.ERLE<- aov(MassLoss~ Plot, data=T3.ERLE)
summary(plot.grass.ERLE)

plot.grass.DICA<- aov(MassLoss~ Plot, data=T3.DICA)
summary(plot.grass.DICA)
```

