---
title: "MonsoonSoilGerminationExperiment"
author: "Taylor A Portman"
date: "2022-10-01"
output: html_document
---
Monsoon soil germination experiment
Taylor Portman and Robin Bradley  

Soil collected from Santa Rita Experimental Range (SRER)  
Soils were collected from 4 plot types
IO: Invasive Open (ERLE dominant)  
IC: Invasive Canopy (ERLE dominat)  
NO: Native Open (DICA dominant)  
NC: Native Canopy (DICA dominant)  

Surface steralized eeds of the invasive grass (ERLE) and four native grasses were plated onto soils of each plot type under two treatments (1) live soil, and (2) sterile soil where soils were tripple autoclaved to remove microbes from the soil.

```{r}
library(tidyverse)
library(ggplot2)
setwd("/Volumes/GoogleDrive/My Drive/projects/ERLE_fungal_symbionts/R")
dat<-read.csv("/Volumes/GoogleDrive/My Drive/projects/ERLE_fungal_symbionts/R/data/Bradley, Portman Live Soil Germination Experiment Datasheet - DataForR.csv")
```


```{r}
head(dat)
germ.dat<-dat%>%
  na.omit()

means <- germ.dat %>% 
  dplyr::group_by(Grass, Plot, SoilType, .add=TRUE) %>%
  dplyr::summarise(
    means=mean(DAY8))
show(means)
```

```{r}
#1. Live vs sterile by seed species
ggplot(germ.dat, aes(Plot, DAY8))+
  geom_boxplot(aes(fill=SoilType))+
  #geom_point(aes(color=SoilType))+
  #facet_wrap(~Grass)+
  scale_fill_manual(values=c("coral", "gray99"))+
  scale_color_manual(values=c("coral", "black"))+
  theme_classic()

ggplot(germ.dat, aes(Plot, DAY8))+
  geom_boxplot(aes(fill=SoilType))+
  #geom_point(aes(color=SoilType))+
  facet_wrap(~Grass)+
  scale_fill_manual(values=c("coral", "gray99"))+
  scale_color_manual(values=c("coral", "black"))+
  theme_classic()

#2. Live vs sterile by plot type
ggplot(germ.dat, aes(Grass, DAY8))+
  geom_boxplot(aes(fill=SoilType))+
  facet_wrap(~Plot)+
  scale_fill_manual(values=c("coral", "gray99"))+
  theme_classic()

ggplot(germ.dat, aes(Grass, DAY8))+
  geom_boxplot(aes(fill=SoilType))+
 # facet_wrap(~Plot)+
  scale_fill_manual(values=c("coral", "gray99"))+
  theme_classic()

#3. Seed species by plot type
ggplot(germ.dat, aes(Plot, DAY8))+
  geom_boxplot(aes(fill=Grass))+
  facet_wrap(~SoilType)+
  theme_classic()

ggplot(germ.dat, aes(Plot, DAY8))+
  geom_boxplot(aes(fill=Grass))+
 # facet_wrap(~SoilType)+
  theme_classic()

#4. Native vs invasive; Live vs sterile
ggplot(germ.dat, aes(Grass, DAY8))+
  geom_boxplot(aes(fill=GrassType))+
  facet_wrap(~SoilType)+
  scale_fill_manual(values=c("goldenrod1", "dodgerblue"))+
  theme_classic()

ggplot(germ.dat, aes(Grass, DAY8))+
  geom_boxplot(aes(fill=GrassType))+
  #facet_wrap(~SoilType)+
  scale_fill_manual(values=c("goldenrod1", "dodgerblue"))+
  theme_classic()

#5. Invasive vs native by plot type
ggplot(germ.dat, aes(Grass, DAY8))+
  geom_boxplot(aes(fill=GrassType))+
  facet_wrap(~SoilType* Cover)+
  scale_fill_manual(values=c("goldenrod1", "dodgerblue"))+
  theme_classic()

ggplot(germ.dat, aes(Grass, DAY8))+
  geom_boxplot(aes(fill=GrassType))+
  #facet_wrap(~SoilType* Cover)+
  scale_fill_manual(values=c("goldenrod1", "dodgerblue"))+
  theme_classic()


#6. Live vs sterile; Canopy vs open
ggplot(germ.dat, aes(Grass, DAY8))+
  geom_boxplot(aes(fill= Cover))+
  facet_wrap(~SoilType)+
  scale_fill_manual(values=c("darkolivegreen", "lavender"))+
  theme_classic()

ggplot(germ.dat, aes(Grass, DAY8))+
  geom_boxplot(aes(fill= Cover))+
 # facet_wrap(~SoilType)+
  scale_fill_manual(values=c("darkolivegreen", "lavender"))+
  theme_classic()

```

